% ══════════════════════════════════════════════════════════════════════════════
%  ObjectIR v2 — Language Specification
% ══════════════════════════════════════════════════════════════════════════════
\documentclass[11pt,a4paper]{article}

% ── Packages ──────────────────────────────────────────────────────────────────
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{parskip}
\usepackage{microtype}
\usepackage{lmodern}
\usepackage{tabularx}
\usepackage{mdframed}
\usepackage{amsmath}

\geometry{margin=2.4cm}

% ── Colors ────────────────────────────────────────────────────────────────────
\definecolor{codebg}{HTML}{F5F5F5}
\definecolor{codeborder}{HTML}{CCCCCC}
\definecolor{kw}{HTML}{0000CC}
\definecolor{str}{HTML}{008800}
\definecolor{comment}{HTML}{888888}
\definecolor{typecolor}{HTML}{7B0000}
\definecolor{linkcolor}{HTML}{0066CC}
\definecolor{sectionrule}{HTML}{3A6EA5}
\definecolor{notebg}{HTML}{FFF8DC}
\definecolor{noteborder}{HTML}{DAA520}
\definecolor{warnbg}{HTML}{FFF0F0}
\definecolor{warnborder}{HTML}{CC4444}
\definecolor{infobg}{HTML}{EEF4FF}
\definecolor{infoborder}{HTML}{5577CC}

% ── Hyperref ──────────────────────────────────────────────────────────────────
\hypersetup{
  colorlinks=true,
  linkcolor=linkcolor,
  urlcolor=linkcolor,
  citecolor=linkcolor,
  pdftitle={ObjectIR v2 --- Language Specification},
  pdfauthor={ObjectIR Project},
}

% ── Code listing styles ───────────────────────────────────────────────────────
\lstdefinelanguage{TextIR}{
  morekeywords={module,version,class,interface,struct,enum,field,method,
                constructor,local,public,private,protected,internal,static,
                virtual,override,abstract,sealed,implements,else,catch,finally,
                true,false,null},
  morestring=[b]",
  morecomment=[l]{//},
  sensitive=true
}

\lstdefinestyle{textir}{
  language=TextIR,
  backgroundcolor=\color{codebg},
  frame=single,
  rulecolor=\color{codeborder},
  basicstyle=\ttfamily\small,
  keywordstyle=\bfseries\color{kw},
  stringstyle=\color{str},
  commentstyle=\itshape\color{comment},
  breaklines=true,
  showstringspaces=false,
  tabsize=4,
  numbers=left,
  numberstyle=\tiny\color{comment},
  numbersep=6pt,
  xleftmargin=1.8em,
}

\lstdefinestyle{pseudo}{
  backgroundcolor=\color{codebg},
  frame=single,
  rulecolor=\color{codeborder},
  basicstyle=\ttfamily\small,
  breaklines=true,
  showstringspaces=false,
  tabsize=4,
  numbers=none,
}

\lstdefinestyle{json}{
  backgroundcolor=\color{codebg},
  frame=single,
  rulecolor=\color{codeborder},
  basicstyle=\ttfamily\small,
  stringstyle=\color{str},
  breaklines=true,
  showstringspaces=false,
  tabsize=2,
  numbers=none,
}

\newcommand{\code}[1]{\texttt{\small #1}}
\newcommand{\typename}[1]{\textcolor{typecolor}{\texttt{\small #1}}}
\newcommand{\kword}[1]{\textbf{\texttt{\small #1}}}
\newcommand{\opcode}[1]{\textbf{\texttt{#1}}}
\newcommand{\nterm}[1]{\textit{$\langle$#1$\rangle$}}
\newcommand{\term}[1]{\texttt{#1}}

% ── Callout boxes ─────────────────────────────────────────────────────────────
\newmdenv[
  backgroundcolor=notebg, linecolor=noteborder, linewidth=1pt,
  innerleftmargin=8pt, innerrightmargin=8pt,
  innertopmargin=6pt, innerbottommargin=6pt,
  skipabove=6pt, skipbelow=6pt,
]{notebox}

\newmdenv[
  backgroundcolor=warnbg, linecolor=warnborder, linewidth=1pt,
  innerleftmargin=8pt, innerrightmargin=8pt,
  innertopmargin=6pt, innerbottommargin=6pt,
  skipabove=6pt, skipbelow=6pt,
]{warnbox}

\newmdenv[
  backgroundcolor=infobg, linecolor=infoborder, linewidth=1pt,
  innerleftmargin=8pt, innerrightmargin=8pt,
  innertopmargin=6pt, innerbottommargin=6pt,
  skipabove=6pt, skipbelow=6pt,
]{infobox}

% ── Headers / footers ─────────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0.4pt}
\fancyhead[L]{\small\textit{ObjectIR v2 --- Language Specification}}
\fancyhead[R]{\small\textit{ObjectIR Project}}
\fancyfoot[C]{\thepage}

% ── Section formatting ────────────────────────────────────────────────────────
\titleformat{\section}{\large\bfseries\color{sectionrule}}{}{0em}{}[{\color{sectionrule}\titlerule[0.8pt]}]
\titleformat{\subsection}{\normalsize\bfseries}{}{0em}{}
\titleformat{\subsubsection}{\normalsize\itshape\bfseries}{}{0em}{}

% ══════════════════════════════════════════════════════════════════════════════
\begin{document}

\begin{titlepage}
  \centering
  \vspace*{3cm}
  {\Huge\bfseries ObjectIR}\\[0.5em]
  {\large\bfseries Version 2}\\[0.8em]
  {\Large Language Specification}\\[2.5em]
  \rule{0.55\linewidth}{0.4pt}\\[1.2em]
  {\normalsize
    A stack-based intermediate representation\\[0.3em]
    and virtual machine definition
  }\\[2.5em]
  \begin{tabular}{ll}
    \textbf{Source format:}   & TextIR (human-readable), JSON, FOB/IR v3 (binary) \\[0.2em]
    \textbf{Execution model:} & Stack machine, one evaluation stack per call frame \\[0.2em]
    \textbf{Type model:}      & Untyped stack; lazy coercion at instruction boundaries \\[0.2em]
    \textbf{Revision:}        & \today \\
  \end{tabular}
  \vfill
  {\small This document is the normative specification of the ObjectIR language:
  its source format, module structure, type system, and instruction set.
  Implementation-specific notes are confined to appendices.}
\end{titlepage}

\tableofcontents
\newpage

% ══════════════════════════════════════════════════════════════════════════════
\section{Introduction}
% ══════════════════════════════════════════════════════════════════════════════

\textbf{ObjectIR} is a typed, object-oriented intermediate representation (IR)
designed to be simple to parse, simple to execute, and straightforward to
target from higher-level languages or toolchains.

A program is expressed as one or more \emph{modules}, each containing class
and interface definitions whose methods are composed of explicit stack-machine
instructions. The IR sits at the same conceptual level as JVM bytecode or
CIL, but its primary surface format (\emph{TextIR}) is human-readable and
hand-writable.

\subsection{Design Goals}

\begin{itemize}[leftmargin=1.8em]
  \item \textbf{Human-readable source.} TextIR is the canonical format: a
    whitespace-significant, keyword-driven text form that lowers to the same
    module representation as the JSON and binary forms.
  \item \textbf{Multiple input formats.} A module may be delivered as TextIR
    source, a JSON object graph, or a compact FOB/IR v3 binary. All three
    forms are semantically equivalent.
  \item \textbf{Embeddable.} A conforming runtime must allow host code to
    load a module, invoke individual methods, and register external (native)
    type and method bindings without touching a parser.
  \item \textbf{Lazy coercion.} Values on the evaluation stack carry no
    static type tag. Numeric widening and other conversions are performed
    at instruction boundaries according to well-defined rules.
  \item \textbf{Structured control flow.} Loops, conditionals, and exception
    handlers are first-class instructions, not jump targets.
\end{itemize}

\subsection{Document Conventions}

\begin{itemize}[leftmargin=1.8em]
  \item \textbf{Normative} text uses \textbf{shall} / \textbf{must}.
    \textbf{Informative} text uses \textit{should} / \textit{may}.
  \item Grammar non-terminals are written \nterm{like-this}; terminals are
    written \term{like-this}.
  \item Instruction mnemonics are typeset \opcode{like-this}.
  \item Type names are typeset \typename{like-this}.
\end{itemize}

% ══════════════════════════════════════════════════════════════════════════════
\section{Module Structure}
\label{sec:module}
% ══════════════════════════════════════════════════════════════════════════════

A \emph{module} is the top-level compilation and loading unit of ObjectIR.

\subsection{Module Contents}

A well-formed module contains:

\begin{itemize}[leftmargin=1.8em]
  \item A \textbf{name}: a non-empty identifier string.
  \item A \textbf{version}: a three-part numeric tuple
    (\textit{major}.\textit{minor}.\textit{patch}), where \textit{patch} is
    optional and defaults to \texttt{0}.
  \item An ordered list of \textbf{type definitions}: classes, interfaces,
    structs, and enums.
  \item A (reserved) \textbf{free-function table}, intended for future
    module-level functions outside any type. Runtimes must accept and ignore
    this field if present.
\end{itemize}

\subsection{Module Identity}

Two modules are considered the \emph{same module} if and only if their names
are identical. Version numbers are informational; the runtime does not perform
version enforcement unless the host explicitly requires it.

\subsection{Active Module}

A runtime instance holds \textbf{exactly one} active module at a time.
Loading a new module replaces the previous one. A \emph{full reset} also
clears all static field state; a \emph{module-only reload} replaces the type
table while optionally preserving heap state.

% ══════════════════════════════════════════════════════════════════════════════
\section{Source Formats}
\label{sec:formats}
% ══════════════════════════════════════════════════════════════════════════════

ObjectIR defines three equivalent surface representations of a module. All
three must express identical semantics; a conforming tool or runtime may
accept any subset.

\begin{longtable}{lp{10cm}}
  \toprule
  \textbf{Format} & \textbf{Description} \\
  \midrule
  \endhead
  \textbf{TextIR}    &
    Human-readable source text (see Section~\ref{sec:textir}).
    The canonical authoring format. A conforming parser produces a module
    representation equivalent to the JSON form. \\[0.4em]
  \textbf{JSON}      &
    A JSON object whose structure mirrors the module schema
    (Section~\ref{sec:schema}). Property names are camelCase.
    This is the recommended interchange format between tools. \\[0.4em]
  \textbf{FOB/IR v3} &
    A compact binary encoding of the module schema
    (Section~\ref{sec:fob}). Intended for distribution and fast loading. \\
  \bottomrule
\end{longtable}

\begin{notebox}
\textbf{Auto-detection hint:} A string whose first non-whitespace character
is \texttt{\{} or \texttt{[} should be treated as JSON; all other strings
should be treated as TextIR. Implementations may expose this heuristic as
an \texttt{Auto} format option.
\end{notebox}

% ══════════════════════════════════════════════════════════════════════════════
\section{TextIR Syntax}
\label{sec:textir}
% ══════════════════════════════════════════════════════════════════════════════

TextIR is a line-oriented, brace-delimited language. Whitespace (outside
string literals) is insignificant except as a token separator. Comments
begin with \term{//} and extend to the end of the line.

\subsection{Module Header}

Every TextIR file must begin with a module header:

\begin{lstlisting}[style=textir]
module <name> version <major>.<minor>[.<patch>]
\end{lstlisting}

\subsection{Type Declarations}

\begin{center}
\begin{tabular}{lll}
  \toprule
  \textbf{Keyword}   & \textbf{Kind}      & \textbf{Notes} \\
  \midrule
  \term{class}     & Reference type   & Instances are heap-allocated objects. \\
  \term{interface} & Contract type    & Declares virtual methods; no fields. \\
  \term{struct}    & Value type       & Semantics are host-defined. \\
  \term{enum}      & Enumeration type & Members are named integer constants. \\
  \bottomrule
\end{tabular}
\end{center}

Any combination of the following \emph{access modifiers} may precede the
type keyword: \term{public}, \term{private}, \term{protected},
\term{internal}.

The following \emph{type modifiers} may also precede the keyword:
\term{abstract}, \term{sealed}.

Inheritance and interface implementation are expressed with \term{:}:

\begin{lstlisting}[style=textir]
class Animal {
    field name: string
    virtual method Speak() -> void { ret }
}

class Dog : Animal {
    override method Speak() -> void implements Animal.Speak {
        ldstr "Woof!"
        call System.Console.WriteLine(string) -> void
        ret
    }
}
\end{lstlisting}

\subsection{Member Declarations}

\begin{lstlisting}[style=textir]
// Field
[static] field <name>: <type>

// Method
[static] [virtual] [override] [abstract]
method <name>(<params>) -> <rettype> [implements <Interface>.<MethodName>]
{
    [local <name>: <type>]*
    <instructions>
}

// Constructor
constructor(<params>) {
    [local <name>: <type>]*
    <instructions>
}
\end{lstlisting}

Parameter lists and local variable declarations share the syntax
\code{name: type}, comma-separated.

\subsection{Grammar Summary}

\begin{lstlisting}[style=pseudo]
module-decl  ::= "module" name "version" version-num
version-num  ::= int "." int ["." int]

type-decl    ::= access-mod* type-mod* type-kw name [":" base-name]
                 "{" member* "}"
type-kw      ::= "class" | "interface" | "struct" | "enum"
access-mod   ::= "public" | "private" | "protected" | "internal"
type-mod     ::= "abstract" | "sealed"

member       ::= field-decl | method-decl | ctor-decl
field-decl   ::= ["static"] "field" name ":" type-name
method-decl  ::= method-mod* "method" name "(" params ")" "->" type-name
                 ["implements" qualified-name] "{" body "}"
method-mod   ::= "static" | "virtual" | "override" | "abstract"
ctor-decl    ::= "constructor" "(" params ")" "{" body "}"

params       ::= (name ":" type-name ("," name ":" type-name)*)?
body         ::= ("local" name ":" type-name)* instruction*
instruction  ::= opcode [operand]
\end{lstlisting}

\subsection{Attributes (Annotations)}

Custom attributes may be applied to types, fields, and methods by placing
\term{@AttributeName} (with an optional positional argument list) immediately
before the declaration:

\begin{lstlisting}[style=textir]
@Serializable
class PlayerData {
    @NonSerialized
    field cachedHash: int32

    @Deprecated("Use SaveAsync instead")
    method Save() -> void { ret }
}
\end{lstlisting}

A runtime must attach attribute metadata to the corresponding declaration;
how it is stored and queried is implementation-defined.

\subsection{String Escape Sequences}

Within string literals the following escape sequences are defined:

\begin{center}
\begin{tabular}{ll}
  \toprule
  \textbf{Sequence} & \textbf{Meaning} \\
  \midrule
  \texttt{\textbackslash n}              & Line feed (U+000A) \\
  \texttt{\textbackslash t}              & Horizontal tab (U+0009) \\
  \texttt{\textbackslash r}              & Carriage return (U+000D) \\
  \texttt{\textbackslash\textbackslash}  & Backslash (U+005C) \\
  \texttt{\textbackslash"}              & Double quote (U+0022) \\
  \bottomrule
\end{tabular}
\end{center}

% ══════════════════════════════════════════════════════════════════════════════
\section{Type System}
\label{sec:types}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Primitive Types}

ObjectIR defines a fixed set of primitive types. Type names in operands are
\textbf{case-insensitive}: the runtime normalises them to lower-case before
lookup, so \texttt{System.String}, \texttt{system.string}, and \texttt{string}
are all equivalent.

\begin{longtable}{lll}
  \toprule
  \textbf{Primary IR name(s)} & \textbf{Alias(es)} & \textbf{Width / Semantics} \\
  \midrule
  \endhead
  \texttt{void}                     & \texttt{system.void}     & No value; return type only. \\
  \texttt{bool}                     & \texttt{system.boolean}  & Boolean true/false. \\
  \texttt{int8}                     & \texttt{system.sbyte}    & 8-bit signed integer. \\
  \texttt{uint8}                    & \texttt{system.byte}     & 8-bit unsigned integer. \\
  \texttt{int16}                    & \texttt{system.int16}    & 16-bit signed integer. \\
  \texttt{uint16}                   & \texttt{system.uint16}   & 16-bit unsigned integer. \\
  \texttt{int32}                    & \texttt{system.int32}    & 32-bit signed integer. \\
  \texttt{uint32}                   & \texttt{system.uint32}   & 32-bit unsigned integer. \\
  \texttt{int64}                    & \texttt{system.int64}    & 64-bit signed integer. \\
  \texttt{uint64}                   & \texttt{system.uint64}   & 64-bit unsigned integer. \\
  \texttt{float32}, \texttt{single} & \texttt{system.single}   & 32-bit IEEE 754 float. \\
  \texttt{float64}, \texttt{double} & \texttt{system.double}   & 64-bit IEEE 754 float. \\
  \texttt{char}                     & \texttt{system.char}     & Unicode scalar (UTF-16 unit). \\
  \texttt{string}                   & \texttt{system.string}   & Immutable Unicode character sequence. \\
  \texttt{object}                   & \texttt{system.object}   & Top type; any value. \\
  \texttt{decimal}                  & \texttt{system.decimal}  & High-precision fixed-point decimal. \\
  \texttt{datetime}                 & \texttt{system.datetime} & Date and time instant. \\
  \texttt{timespan}                 & \texttt{system.timespan} & Duration / time interval. \\
  \texttt{guid}                     & \texttt{system.guid}     & 128-bit globally unique identifier. \\
  \bottomrule
\end{longtable}

\subsection{User-Defined Types}

A type defined in the module is either a \textbf{class}, \textbf{interface},
\textbf{struct}, or \textbf{enum}. At runtime, instances of user-defined
types are \emph{managed objects}: heap-allocated records that carry:

\begin{itemize}[leftmargin=1.8em]
  \item A \textbf{type name} identifying the declared type.
  \item A \textbf{field store}: a mapping from field name to value.
  \item An optional \textbf{host backing object} supplied by the host runtime
    (see Section~\ref{sec:hostinterop}).
  \item Zero or more \textbf{attribute instances} attached at load time.
\end{itemize}

\subsection{Inheritance}

A class may declare at most one \emph{base class}. A class or struct may
declare any number of \emph{implemented interfaces}. Runtimes are not
required to perform inheritance-aware type checks in v2
(see Section~\ref{sec:limits}).

\subsection{Null}

\texttt{null} is a valid value for any non-primitive type. The
\opcode{ldnull} instruction places it on the stack.

\subsection{Value Coercion Semantics}
\label{sec:coerce}

Because the evaluation stack carries untyped values, the runtime applies
the following coercion rules at instruction boundaries:

\begin{longtable}{lp{9.5cm}}
  \toprule
  \textbf{Target type} & \textbf{Rule} \\
  \midrule
  \endhead
  \texttt{bool}      &
    \texttt{null}, integer \texttt{0}, and empty string \texttt{""} are
    \texttt{false}; all other values are \texttt{true}. \\
  integer type       &
    Floating-point values are truncated toward zero. Strings are parsed
    as base-10 integers using the invariant locale. Out-of-range results
    are implementation-defined but must not raise an unhandled exception. \\
  floating-point     &
    Integer values are widened exactly. Strings are parsed as decimal
    floating-point using the invariant locale. \\
  \texttt{string}    &
    Any non-null value is converted to its canonical string representation.
    \texttt{null} converts to \texttt{null}. \\
  \texttt{object}    &
    Identity; any value is valid. \\
  \bottomrule
\end{longtable}

Arithmetic instructions select \emph{float mode} (operands widened to
64-bit float) when either operand is a floating-point value or a string;
otherwise \emph{integer mode} (64-bit signed) is used.

% ══════════════════════════════════════════════════════════════════════════════
\section{Instruction Set Reference}
\label{sec:isa}
% ══════════════════════════════════════════════════════════════════════════════

ObjectIR uses a \textbf{stack machine} execution model. Each method
invocation creates a new \emph{call frame} that owns:

\begin{itemize}[leftmargin=1.8em, noitemsep]
  \item An \textbf{evaluation stack} of untyped values.
  \item A named \textbf{argument table} (index \texttt{0} / name
    \texttt{this} is the receiver for instance methods).
  \item A named \textbf{local variable table}.
  \item A \textbf{pending exception} slot.
  \item A \textbf{program counter} (index into the instruction list).
\end{itemize}

Operands are described using the JSON operand format of
Section~\ref{sec:schema}.

\subsection{Stack Manipulation}

\begin{longtable}{lp{10.5cm}}
  \toprule
  \textbf{Opcode} & \textbf{Description} \\
  \midrule
  \endhead
  \opcode{nop}    & No-operation; advances the program counter only. \\
  \opcode{dup}    & Duplicate the top value on the evaluation stack. \\
  \opcode{pop}    & Discard the top value. \\
  \opcode{ldnull} & Push \texttt{null} onto the stack. \\
  \bottomrule
\end{longtable}

\subsection{Load Constants}

\begin{longtable}{lp{10.5cm}}
  \toprule
  \textbf{Opcode} & \textbf{Description} \\
  \midrule
  \endhead
  \opcode{ldc}    &
    Push a typed constant.
    Operand: \code{\{value: ..., type: "<typename>"\}}.
    \texttt{type} defaults to \texttt{int64}. \\
  \opcode{ldc.i4} & Push a 32-bit integer constant. Operand: bare integer. \\
  \opcode{ldc.i8} & Push a 64-bit integer constant. Operand: bare integer. \\
  \opcode{ldc.r4} & Push a 32-bit float constant. Operand: bare number. \\
  \opcode{ldc.r8} & Push a 64-bit float constant. Operand: bare number. \\
  \opcode{ldstr}  &
    Push a string constant.
    Operand: \code{\{value: "..."\}} or a bare quoted string. \\
  \bottomrule
\end{longtable}

\subsection{Arguments and Local Variables}

\begin{longtable}{lp{10.5cm}}
  \toprule
  \textbf{Opcode} & \textbf{Description} \\
  \midrule
  \endhead
  \opcode{ldarg}  &
    Push a method argument.
    Operand: parameter name or 0-based index.
    Index \texttt{0} / name \texttt{this} is the receiver. \\
  \opcode{starg}  & Pop a value and store it in the named argument slot. \\
  \opcode{ldloc}  & Push a local variable by name. \\
  \opcode{stloc}  & Pop a value and store it in the named local variable. \\
  \bottomrule
\end{longtable}

\subsection{Field Access}

\begin{longtable}{lp{10.5cm}}
  \toprule
  \textbf{Opcode} & \textbf{Description} \\
  \midrule
  \endhead
  \opcode{ldfld}  &
    Pop an object (or use the implicit receiver) and push the value of the
    named instance field. If the object has a host backing, the runtime
    \textit{should} read the corresponding property on the host object
    before consulting the field store. \\
  \opcode{stfld}  &
    Pop a value, then pop an object; store the value in the named instance
    field (and on the host backing, if present). \\
  \opcode{ldsfld} &
    Push the value of a static field.
    Operand: \code{\{field: \{declaringType, name\}\}}. \\
  \opcode{stsfld} & Pop a value and store it in the named static field. \\
  \bottomrule
\end{longtable}

\subsection{Arithmetic}

\begin{longtable}{lp{10.5cm}}
  \toprule
  \textbf{Opcode} & \textbf{Description} \\
  \midrule
  \endhead
  \opcode{add} &
    Pop $b$, pop $a$, push $a + b$.
    In float mode, also performs string concatenation when either operand
    is a string. \\
  \opcode{sub} & Pop $b$, pop $a$, push $a - b$. \\
  \opcode{mul} & Pop $b$, pop $a$, push $a \times b$. \\
  \opcode{div} & Pop $b$, pop $a$, push $a \div b$. \\
  \opcode{rem} & Pop $b$, pop $a$, push $a \bmod b$. \\
  \opcode{neg} & Pop $v$, push $-v$. \\
  \bottomrule
\end{longtable}

Mode selection follows Section~\ref{sec:coerce}: if either operand is
floating-point (or a string for \opcode{add}), both are promoted to 64-bit
float; otherwise 64-bit signed integer arithmetic is used.

\subsection{Logical}

\begin{longtable}{lp{10.5cm}}
  \toprule
  \textbf{Opcode} & \textbf{Description} \\
  \midrule
  \endhead
  \opcode{not} & Pop $v$, push the boolean negation: \texttt{!ToBool($v$)}. \\
  \bottomrule
\end{longtable}

\subsection{Comparison}

Each opcode pops $b$ then $a$ and pushes a \texttt{bool}.

\begin{longtable}{ll}
  \toprule
  \textbf{Opcode} & \textbf{Condition} \\
  \midrule
  \endhead
  \opcode{ceq} & $a = b$ \\
  \opcode{cne} & $a \neq b$ \\
  \opcode{cgt} & $a > b$ \\
  \opcode{cge} & $a \geq b$ \\
  \opcode{clt} & $a < b$ \\
  \opcode{cle} & $a \leq b$ \\
  \bottomrule
\end{longtable}

String comparisons use ordinal (byte-by-byte) ordering.
Boolean values are supported for \opcode{ceq} and \opcode{cne} only.
All other operand pairs are promoted to 64-bit float before comparison.

\subsection{Type Conversion and Testing}

\begin{longtable}{lp{10.5cm}}
  \toprule
  \textbf{Opcode}    & \textbf{Description} \\
  \midrule
  \endhead
  \opcode{conv}      &
    Pop a value and push it converted to the named target type according
    to the coercion rules in Section~\ref{sec:coerce}. \\
  \opcode{castclass} &
    Assert that the top-of-stack managed object has the given type name;
    raise a cast error if it does not.
    (Exact type-name match only in v2; see Section~\ref{sec:limits}.) \\
  \opcode{isinst}    &
    Pop a value; push \texttt{true} if it is a managed object with the
    given type name, otherwise \texttt{false}. \\
  \bottomrule
\end{longtable}

\begin{warnbox}
\textbf{v2 limitation:} \opcode{castclass} and \opcode{isinst} perform
\emph{exact} type-name matching. Walking an inheritance chain is not required
of conforming implementations in this version.
\end{warnbox}

\subsection{Object and Array Creation}

\begin{longtable}{lp{10.5cm}}
  \toprule
  \textbf{Opcode}  & \textbf{Description} \\
  \midrule
  \endhead
  \opcode{newobj}  &
    Create a new managed object of the named type and push it. If the type
    is a registered host type, the runtime \textit{should} also construct
    the corresponding host backing object. \\
  \opcode{newarr}  &
    Push a new, empty, dynamically-typed array. \\
  \opcode{ldelem}  &
    Pop an index (0-based integer), pop an array;
    push \texttt{array[index]}. \\
  \opcode{stelem}  &
    Pop a value, pop an index, pop an array;
    set \texttt{array[index] = value}. \\
  \bottomrule
\end{longtable}

\begin{notebox}
Arrays in ObjectIR v2 are untyped and variable-length. Typed array variants
(e.g.\ \texttt{int[]}) are not defined in this version.
\end{notebox}

\subsection{Method Calls}

\begin{longtable}{lp{10.5cm}}
  \toprule
  \textbf{Opcode}   & \textbf{Description} \\
  \midrule
  \endhead
  \opcode{call}     &
    Pop arguments left-to-right, then pop the receiver (for instance
    methods); invoke the target method.
    Operand: \code{\{method: \{declaringType, name, returnType,
    parameterTypes[]\}\}}. \\
  \opcode{callvirt} &
    Same as \opcode{call}, but dispatch is virtual: the actual type of the
    receiver determines which override is executed. \\
  \opcode{ret}      &
    Return from the current frame. If a value is present on the evaluation
    stack it is transferred to the caller's stack. \\
  \bottomrule
\end{longtable}

\subsubsection{Call Resolution Order}
\label{sec:callres}

A conforming runtime must attempt the following sequence for every
\opcode{call} and \opcode{callvirt}:

\begin{enumerate}[leftmargin=2em]
  \item \textbf{Exact native/host lookup.} Match the full signature string
    \code{DeclaringType.Name(param0,param1,...)} against the registered
    host method table.
  \item \textbf{Simplified native/host lookup.} Retry with parameter type
    names reduced to their unqualified, lower-case short names
    (e.g.\ \code{System.String} $\to$ \code{string}).
  \item \textbf{Host reflection.} If the runtime supports reflective dispatch
    and it is enabled, invoke the method on the host backing object.
  \item \textbf{IR method dispatch.} Look up the method in the loaded
    module's type table.
\end{enumerate}

If none of the steps resolve the call, a \textbf{missing-method} error is
raised.

\subsection{Control Flow}

\begin{longtable}{lp{10.5cm}}
  \toprule
  \textbf{Opcode}    & \textbf{Description} \\
  \midrule
  \endhead
  \opcode{if}        &
    Evaluate the condition operand (Section~\ref{sec:conditions}).
    If truthy, execute \code{thenBlock};
    otherwise execute the optional \code{elseBlock}. \\
  \opcode{while}     &
    Repeatedly evaluate the condition and execute \code{body}.
    A \opcode{break} inside \code{body} exits the loop;
    a \opcode{continue} skips to the next condition evaluation. \\
  \opcode{break}     & Exit the nearest enclosing \opcode{while}. \\
  \opcode{continue}  & Skip to the next iteration of the nearest \opcode{while}. \\
  \opcode{try}       &
    Execute \code{tryBlock}. If an exception is raised, test each
    \code{catchBlock} in order (Section~\ref{sec:exceptions}).
    Execute the optional \code{finallyBlock} in all cases. \\
  \opcode{throw}     &
    Pop a value and raise it as a pending exception on the current frame. \\
  \bottomrule
\end{longtable}

\subsubsection{Condition Kinds}
\label{sec:conditions}

A condition embedded in an \opcode{if} or \opcode{while} operand has a
\texttt{kind} field that determines how it is evaluated:

\begin{longtable}{lp{10cm}}
  \toprule
  \textbf{Kind}       & \textbf{Evaluation} \\
  \midrule
  \endhead
  \code{stack}      &
    Pop the top-of-stack value and coerce it to \texttt{bool}
    (Section~\ref{sec:coerce}).
    An empty stack is treated as \texttt{false}. \\
  \code{binary}     &
    Pop $r$, pop $l$; apply the named comparison opcode
    (\opcode{ceq}, \opcode{cgt}, \ldots) and use the result. \\
  \code{expression} &
    Execute a single embedded instruction and pop the result. \\
  \code{block}      &
    Execute a sequence of instructions and pop the final result. \\
  \bottomrule
\end{longtable}

% ══════════════════════════════════════════════════════════════════════════════
\section{Exception Model}
\label{sec:exceptions}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Raising Exceptions}

The \opcode{throw} instruction accepts any value: a managed object, a host
exception type, or a plain string. The value is stored in the
\emph{pending exception} slot of the current frame and instruction execution
in that frame halts.

\subsection{Structured Exception Handling}

The \opcode{try} instruction provides first-class structured exception
handling. Its operand defines up to three regions:

\begin{itemize}[leftmargin=1.8em]
  \item \code{tryBlock} --- the guarded instructions.
  \item \code{catchBlocks} --- an ordered list of catch clauses, each with
    an optional \code{type} field and a \code{body}.
  \item \code{finallyBlock} --- instructions executed unconditionally after
    the try/catch, regardless of outcome.
\end{itemize}

Catch clauses are matched in declaration order against the pending exception:

\begin{enumerate}[leftmargin=2em]
  \item \textbf{Catch-all} (no \code{type} field) --- matches any value.
  \item \textbf{Type name match} --- the pending exception is a managed object
    whose type name equals the catch clause's \code{type}.
  \item \textbf{Host type match} --- the host runtime may additionally apply
    assignability checks against registered host types.
\end{enumerate}

\subsection{Exception Propagation}

If no catch clause in the current frame matches, the pending exception
propagates to the caller's frame. This continues until either a matching
clause is found or the call stack is exhausted. An unhandled exception at
the top of the stack is delivered to the host via the exception hook.

\subsection{Control-Flow Exception Tokens}

The \opcode{break} and \opcode{continue} instructions are implemented as
\emph{structured control-flow tokens}. They must not be catchable by user
\opcode{try}/\opcode{catch} blocks and must not propagate outside the
enclosing \opcode{while} instruction.

% ══════════════════════════════════════════════════════════════════════════════
\section{Module Schema (JSON / In-Memory)}
\label{sec:schema}
% ══════════════════════════════════════════════════════════════════════════════

This section defines the canonical JSON structure of a module. The same
structure is used as the in-memory representation produced by a TextIR
parser and consumed by a runtime. All property names are camelCase.

\subsection{Top-Level Object}

\begin{lstlisting}[style=json]
{
  "name":      "<module-name>",
  "version":   "<major>.<minor>.<patch>",
  "types":     [ <TypeObject>, ... ],
  "functions": null   // reserved; must be accepted and ignored
}
\end{lstlisting}

\subsection{Type Object}

\begin{lstlisting}[style=json]
{
  "kind":           "class" | "interface" | "struct" | "enum",
  "name":           "<SimpleName>",
  "_namespace":     "<Namespace>" | null,
  "baseType":       "<QualifiedName>" | null,
  "baseInterfaces": [ "<QualifiedName>", ... ],
  "fields":         [ <FieldObject>, ... ],
  "methods":        [ <MethodObject>, ... ]
}
\end{lstlisting}

\subsection{Field Object}

\begin{lstlisting}[style=json]
{
  "name":       "<name>",
  "type":       "<typename>",
  "isStatic":   true | false,
  "attributes": [ <AttributeObject>, ... ]
}
\end{lstlisting}

\subsection{Method Object}

\begin{lstlisting}[style=json]
{
  "name":         "<name>",
  "returnType":   "<typename>",
  "isStatic":     true | false,
  "isVirtual":    true | false,
  "isOverride":   true | false,
  "isAbstract":   true | false,
  "parameters":   [ { "name": "...", "type": "..." }, ... ],
  "locals":       [ { "name": "...", "type": "..." }, ... ],
  "instructions": [ <InstructionObject>, ... ],
  "attributes":   [ <AttributeObject>, ... ]
}
\end{lstlisting}

\subsection{Instruction Object}

\begin{lstlisting}[style=json]
{
  "opCode":  "<mnemonic>",
  "operand": <any JSON value>
}
\end{lstlisting}

The \texttt{operand} is opcode-specific. Common patterns:

\begin{longtable}{lp{8cm}}
  \toprule
  \textbf{Opcode(s)} & \textbf{Operand form} \\
  \midrule
  \endhead
  \opcode{ldstr}                          & Bare string or \code{\{"value":"..."\}} \\
  \opcode{ldc}, \opcode{ldc.*}            & \code{\{"value": ..., "type": "<typename>"\}} or bare number \\
  \opcode{ldarg}, \opcode{starg},
  \opcode{ldloc}, \opcode{stloc}          & Parameter/local name (string) or 0-based index (integer) \\
  \opcode{ldfld}, \opcode{stfld}          & Field name string \\
  \opcode{ldsfld}, \opcode{stsfld}        & \code{\{"field":\{"declaringType":"...","name":"..."\}\}} \\
  \opcode{call}, \opcode{callvirt}        &
    \code{\{"method":\{"declaringType":"...","name":"...",}
    \code{"returnType":"...","parameterTypes":["..."]\}\}} \\
  \opcode{newobj}, \opcode{conv},
  \opcode{castclass}, \opcode{isinst}     & Type name string \\
  \opcode{if}                             &
    \code{\{"condition":<CondObj>,"thenBlock":[...],}
    \code{"elseBlock":[...]\}} \\
  \opcode{while}                          & \code{\{"condition":<CondObj>,"body":[...]\}} \\
  \opcode{try}                            &
    \code{\{"tryBlock":[...],"catchBlocks":[...],}
    \code{"finallyBlock":[...]\}} \\
  \bottomrule
\end{longtable}

\subsection{Condition Object}

\begin{lstlisting}[style=json]
{
  "kind":      "stack" | "binary" | "expression" | "block",
  "operation": "<opcode>",              // for "binary"
  "body":      [ <InstructionObject>, ... ]  // for "expression" / "block"
}
\end{lstlisting}

\subsection{Attribute Object}

\begin{lstlisting}[style=json]
{
  "name": "<AttributeTypeName>",
  "args": [ <value>, ... ]
}
\end{lstlisting}

% ══════════════════════════════════════════════════════════════════════════════
\section{FOB/IR v3 Binary Format}
\label{sec:fob}
% ══════════════════════════════════════════════════════════════════════════════

The FOB/IR v3 format is a compact binary encoding of the module schema
defined in Section~\ref{sec:schema}. It is intended for distribution and
fast loading; the semantics are identical to the JSON and TextIR forms.

\subsection{File Layout}

\begin{longtable}{lp{9cm}}
  \toprule
  \textbf{Section} & \textbf{Contents} \\
  \midrule
  \endhead
  \textbf{Header}        &
    4-byte magic \texttt{0x464F4249} (\texttt{FOBI}), 1-byte format
    version (\texttt{0x03}), module name (length-prefixed UTF-8), version
    triple (three \texttt{uint16}). \\
  \textbf{String pool}   &
    A length-prefixed table of interned UTF-8 strings referenced by index
    throughout the rest of the file. \\
  \textbf{Type table}    &
    Sequential type records: kind byte, name index, namespace index,
    base type index, interface index list, field count + field records,
    method count + method records. \\
  \textbf{Method bodies} &
    Per-method: parameter count + parameter records, local count + local
    records, instruction count + instruction records. \\
  \bottomrule
\end{longtable}

\subsection{Binary Opcode Map}

Each instruction is stored as a 1-byte opcode followed by a variable-length
operand.

\begin{longtable}{llll}
  \toprule
  \textbf{Byte} & \textbf{Mnemonic} & \textbf{Byte} & \textbf{Mnemonic} \\
  \midrule
  \endhead
  \texttt{0x00} & \texttt{nop}       & \texttt{0x10} & \texttt{ldsfld}   \\
  \texttt{0x01} & \texttt{ldc}       & \texttt{0x11} & \texttt{stsfld}   \\
  \texttt{0x02} & \texttt{ldstr}     & \texttt{0x12} & \texttt{newobj}   \\
  \texttt{0x03} & \texttt{ldarg}     & \texttt{0x13} & \texttt{newarr}   \\
  \texttt{0x04} & \texttt{starg}     & \texttt{0x14} & \texttt{ldelem}   \\
  \texttt{0x05} & \texttt{ldloc}     & \texttt{0x15} & \texttt{stelem}   \\
  \texttt{0x06} & \texttt{stloc}     & \texttt{0x16} & \texttt{call}     \\
  \texttt{0x07} & \texttt{add}       & \texttt{0x17} & \texttt{callvirt} \\
  \texttt{0x08} & \texttt{sub}       & \texttt{0x18} & \texttt{ret}      \\
  \texttt{0x09} & \texttt{mul}       & \texttt{0x19} & \texttt{if}       \\
  \texttt{0x0A} & \texttt{div}       & \texttt{0x1A} & \texttt{while}    \\
  \texttt{0x0B} & \texttt{rem}       & \texttt{0x1B} & \texttt{break}    \\
  \texttt{0x0C} & \texttt{neg}       & \texttt{0x1C} & \texttt{continue} \\
  \texttt{0x0D} & \texttt{ceq}       & \texttt{0x1D} & \texttt{try}      \\
  \texttt{0x0E} & \texttt{cne}       & \texttt{0x1E} & \texttt{throw}    \\
  \texttt{0x0F} & \texttt{ldfld}     & \texttt{0x1F} & \texttt{conv}     \\
  \texttt{0x20} & \texttt{castclass} & \texttt{0x21} & \texttt{isinst}   \\
  \texttt{0x22} & \texttt{dup}       & \texttt{0x23} & \texttt{pop}      \\
  \texttt{0x24} & \texttt{ldnull}    & \texttt{0x25} & \texttt{not}      \\
  \texttt{0x26} & \texttt{cgt}       & \texttt{0x27} & \texttt{cge}      \\
  \texttt{0x28} & \texttt{clt}       & \texttt{0x29} & \texttt{cle}      \\
  \texttt{0x2A} & \texttt{stfld}     & \texttt{0x2B} & \texttt{ldc.i4}   \\
  \texttt{0x2C} & \texttt{ldc.i8}    & \texttt{0x2D} & \texttt{ldc.r4}   \\
  \texttt{0x2E} & \texttt{ldc.r8}    &               &                   \\
  \bottomrule
\end{longtable}

% ══════════════════════════════════════════════════════════════════════════════
\section{Host Interoperability}
\label{sec:hostinterop}
% ══════════════════════════════════════════════════════════════════════════════

This section defines requirements for runtimes that allow host code to
expose native types and methods to IR programs. All features in this section
are \textbf{optional}; a minimal runtime that only executes pure IR code
need not implement them.

\subsection{Host Type Registration}

A runtime \textit{may} allow the host to register \emph{external types}
under IR type names. Once registered:

\begin{itemize}[leftmargin=1.8em]
  \item \opcode{newobj} for that type name constructs a host instance and
    attaches it as the backing of the managed object wrapper.
  \item \opcode{ldfld} / \opcode{stfld} consult the host object's property
    store before the managed object's field dictionary.
  \item \opcode{ldsfld} / \opcode{stsfld} may access static members on the
    host type.
\end{itemize}

\subsection{Native Method Bindings}

A runtime \textit{may} allow the host to register \emph{native method
bindings}: callbacks keyed by a call signature string of the form:

\[
\texttt{DeclaringType.MethodName(param0Type,param1Type,...)}
\]

When the runtime resolves a \opcode{call} or \opcode{callvirt}, it checks
the native binding table before performing IR method lookup (steps 1--3 of
Section~\ref{sec:callres}).

A native binding callback receives:
\begin{itemize}[leftmargin=1.8em, noitemsep]
  \item The receiver object (or a null equivalent for static calls).
  \item A positional list of argument values in declaration order.
\end{itemize}

It returns a single result value (or a null equivalent for void methods).

\subsection{Observability Hooks}

A runtime \textit{should} provide the following hooks for diagnostics and
tooling:

\begin{itemize}[leftmargin=1.8em]
  \item \textbf{Step hook} --- invoked before each instruction with the
    current frame and instruction as arguments. Useful for debuggers,
    profilers, and instruction-level traces.
  \item \textbf{Exception hook} --- invoked when an exception escapes the
    top of the call stack.
\end{itemize}

% ══════════════════════════════════════════════════════════════════════════════
\section{Runtime Conformance Requirements}
\label{sec:conformance}
% ══════════════════════════════════════════════════════════════════════════════

A conforming ObjectIR v2 runtime \textbf{must}:

\begin{enumerate}[leftmargin=2em]
  \item Accept at least one of the three module formats
    (TextIR, JSON, FOB/IR v3) as input.
  \item Maintain a call stack of frames with independent evaluation stacks.
  \item Implement all opcodes defined in Section~\ref{sec:isa}.
  \item Apply the coercion rules of Section~\ref{sec:coerce} at instruction
    boundaries.
  \item Implement the call resolution order of Section~\ref{sec:callres}.
  \item Implement structured exception handling as defined in
    Section~\ref{sec:exceptions}, including the requirement that
    \opcode{break}/\opcode{continue} tokens are not user-catchable.
  \item Expose an entry-point invocation mechanism (canonically targeting a
    static \texttt{Program.Main} with no parameters).
  \item Expose a named method invocation mechanism allowing the host to call
    any method by qualified name with arbitrary arguments.
  \item Expose a named method existence check that inspects module metadata
    without executing code.
\end{enumerate}

A conforming runtime \textit{should} additionally:

\begin{enumerate}[leftmargin=2em, resume]
  \item Implement host type registration (Section~\ref{sec:hostinterop}).
  \item Implement native method binding (Section~\ref{sec:hostinterop}).
  \item Implement step and exception observability hooks.
  \item Pre-register the built-in \texttt{System.Console} and
    \texttt{System.String} native bindings listed in
    Appendix~\ref{app:builtins}.
\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section{Worked Examples}
% ══════════════════════════════════════════════════════════════════════════════

\subsection{Hello World}

\begin{lstlisting}[style=textir, caption={Hello World in TextIR}]
module HelloWorld version 1.0.0

class Program {
    static method Main() -> void {
        ldstr "Hello, ObjectIR!"
        call System.Console.WriteLine(string) -> void
        ret
    }
}
\end{lstlisting}

\subsection{Fibonacci (Recursive)}

\begin{lstlisting}[style=textir, caption={Recursive Fibonacci}]
module Fibonacci version 1.0.0

class Program {
    static method Fib(n: int32) -> int32 {
        ldarg n
        ldc.i4 2
        clt
        if { "kind": "stack",
             "thenBlock": [
               { "opCode": "ldarg", "operand": "n" },
               { "opCode": "ret" }
             ] }

        ldarg n
        ldc.i4 1
        sub
        call Program.Fib(int32) -> int32

        ldarg n
        ldc.i4 2
        sub
        call Program.Fib(int32) -> int32

        add
        ret
    }

    static method Main() -> void {
        ldc.i4 10
        call Program.Fib(int32) -> int32
        call System.Console.WriteLine(int32) -> void
        ret
    }
}
\end{lstlisting}

\subsection{Instance Classes and Fields}

\begin{lstlisting}[style=textir, caption={Instance class with constructor}]
module ClassDemo version 1.0.0

class Vector2D {
    field x: float32
    field y: float32

    constructor(x: float32, y: float32) {
        ldarg this
        ldarg x
        stfld x
        ldarg this
        ldarg y
        stfld y
        ret
    }

    // returns sqrt(x*x + y*y)
    method Length() -> float32 {
        ldarg this
        ldfld x
        ldarg this
        ldfld x
        mul
        ldarg this
        ldfld y
        ldarg this
        ldfld y
        mul
        add
        call System.Math.Sqrt(float64) -> float64
        conv float32
        ret
    }
}

class Program {
    static method Main() -> void {
        ldc.r4 3.0
        ldc.r4 4.0
        newobj Vector2D
        callvirt Vector2D.Length() -> float32
        call System.Console.WriteLine(object) -> void
        ret
    }
}
\end{lstlisting}

\subsection{Exception Handling}

\begin{lstlisting}[style=textir, caption={try / catch / finally}]
module ExceptionDemo version 1.0.0

class MyError {
    field message: string
    constructor(message: string) {
        ldarg this
        ldarg message
        stfld message
        ret
    }
}

class Program {
    static method Main() -> void {
        try {
            "tryBlock": [
                { "opCode": "newobj",  "operand": "MyError" },
                { "opCode": "throw" }
            ],
            "catchBlocks": [ {
                "type": "MyError",
                "variable": "err",
                "body": [
                    { "opCode": "ldstr", "operand": "caught!" },
                    { "opCode": "call",
                      "operand": { "method": {
                        "declaringType": "System.Console",
                        "name": "WriteLine",
                        "returnType": "void",
                        "parameterTypes": ["string"] } } }
                ]
            } ],
            "finallyBlock": [
                { "opCode": "ldstr", "operand": "always runs" },
                { "opCode": "call",
                  "operand": { "method": {
                    "declaringType": "System.Console",
                    "name": "WriteLine",
                    "returnType": "void",
                    "parameterTypes": ["string"] } } }
            ]
        }
        ret
    }
}
\end{lstlisting}

% ══════════════════════════════════════════════════════════════════════════════
\section{Known Limitations (v2)}
\label{sec:limits}
% ══════════════════════════════════════════════════════════════════════════════

The following features are explicitly \textbf{out of scope} for ObjectIR v2
and are reserved for future versions:

\begin{itemize}[leftmargin=1.8em]
  \item \textbf{Generics.} Generic type parameters and instantiation are not
    defined. Generic host APIs must be exposed through native method bindings.
  \item \textbf{Typed arrays.} All arrays are untyped. Distinct
    \texttt{int[]}, \texttt{float[]}, etc.\ are not distinguished.
  \item \textbf{Inheritance-aware type tests.} \opcode{castclass} and
    \opcode{isinst} match by exact type name only.
  \item \textbf{Method overloading on IR-defined types.} Only one method
    with a given name is permitted per IR type in v2; overloading resolution
    applies only to host native bindings.
  \item \textbf{Module-level free functions.} The \texttt{functions} field is
    reserved; runtimes must accept it but need not execute it.
  \item \textbf{Concurrent execution.} The specification does not define
    thread-safety semantics for concurrent use of a single runtime instance.
\end{itemize}

% ══════════════════════════════════════════════════════════════════════════════
\appendix
% ══════════════════════════════════════════════════════════════════════════════

\section{Recommended Built-in Native Method Bindings}
\label{app:builtins}

A conforming runtime \textit{should} pre-register the following native
bindings. The table describes the expected observable behaviour; the
implementation mechanism is host-defined.

\begin{longtable}{lp{7.5cm}}
  \toprule
  \textbf{Signature} & \textbf{Expected behaviour} \\
  \midrule
  \endhead
  \texttt{System.Console.WriteLine(string)}
    & Write the argument followed by a newline to standard output. \\
  \texttt{System.Console.WriteLine(int32)}
    & Same, with an integer argument. \\
  \texttt{System.Console.WriteLine(object)}
    & Same, converting the argument to string first. \\
  \texttt{System.Console.WriteLine(string,string)}
    & Concatenate the two arguments and write a line. \\
  \texttt{System.Console.WriteLine(string,object)}
    & Format the first argument as a template with the second; write a line. \\
  \texttt{System.Console.Write(string)}
    & Write the argument to standard output without a newline. \\
  \texttt{System.Console.Write(string,string)}
    & Concatenate the two arguments and write without a newline. \\
  \texttt{System.Console.ReadLine()}
    & Read one line from standard input and return it as a string. \\
  \texttt{System.Console.Clear()}
    & Clear the standard output terminal if supported; otherwise no-op. \\
  \texttt{System.String.Concat(string,string)}
    & Return the concatenation of the two string arguments. \\
  \texttt{System.String.Format(string,object)}
    & Return the first argument as a format template applied to the second. \\
  \texttt{System.String.Format(string,object,object)}
    & Same with two substitution arguments. \\
  \texttt{System.String.IsNullOrEmpty(string)}
    & Return \texttt{true} if the argument is \texttt{null} or \texttt{""}. \\
  \texttt{System.Object.ToString()}
    & Return the string representation of the receiver. \\
  \texttt{System.Object.GetType()}
    & Return a description of the receiver's runtime type. \\
  \bottomrule
\end{longtable}

\section{Opcode Quick Reference}
\label{app:opcodes}

\begin{longtable}{lll}
  \toprule
  \textbf{Mnemonic} & \textbf{Category} & \textbf{Stack effect} \\
  \midrule
  \endhead
  \opcode{nop}       & Stack      & --- \\
  \opcode{dup}       & Stack      & $v \to v,\, v$ \\
  \opcode{pop}       & Stack      & $v \to$ \\
  \opcode{ldnull}    & Stack      & $\to \texttt{null}$ \\
  \opcode{ldc}       & Constants  & $\to c$ \\
  \opcode{ldc.i4}    & Constants  & $\to c_{\text{i32}}$ \\
  \opcode{ldc.i8}    & Constants  & $\to c_{\text{i64}}$ \\
  \opcode{ldc.r4}    & Constants  & $\to c_{\text{f32}}$ \\
  \opcode{ldc.r8}    & Constants  & $\to c_{\text{f64}}$ \\
  \opcode{ldstr}     & Constants  & $\to s$ \\
  \opcode{ldarg}     & Variables  & $\to a$ \\
  \opcode{starg}     & Variables  & $v \to$ \\
  \opcode{ldloc}     & Variables  & $\to l$ \\
  \opcode{stloc}     & Variables  & $v \to$ \\
  \opcode{ldfld}     & Fields     & $\mathit{obj} \to v$ \\
  \opcode{stfld}     & Fields     & $\mathit{obj}, v \to$ \\
  \opcode{ldsfld}    & Fields     & $\to v$ \\
  \opcode{stsfld}    & Fields     & $v \to$ \\
  \opcode{add}       & Arithmetic & $a, b \to a{+}b$ \\
  \opcode{sub}       & Arithmetic & $a, b \to a{-}b$ \\
  \opcode{mul}       & Arithmetic & $a, b \to a{\times}b$ \\
  \opcode{div}       & Arithmetic & $a, b \to a{\div}b$ \\
  \opcode{rem}       & Arithmetic & $a, b \to a \bmod b$ \\
  \opcode{neg}       & Arithmetic & $v \to {-}v$ \\
  \opcode{not}       & Logical    & $v \to \neg v$ \\
  \opcode{ceq}       & Comparison & $a, b \to a{=}b$ \\
  \opcode{cne}       & Comparison & $a, b \to a{\neq}b$ \\
  \opcode{cgt}       & Comparison & $a, b \to a{>}b$ \\
  \opcode{cge}       & Comparison & $a, b \to a{\geq}b$ \\
  \opcode{clt}       & Comparison & $a, b \to a{<}b$ \\
  \opcode{cle}       & Comparison & $a, b \to a{\leq}b$ \\
  \opcode{conv}      & Type       & $v \to v'$ \\
  \opcode{castclass} & Type       & $v \to v$ (or error) \\
  \opcode{isinst}    & Type       & $v \to \texttt{bool}$ \\
  \opcode{newobj}    & Objects    & $\to \mathit{obj}$ \\
  \opcode{newarr}    & Objects    & $\to \mathit{arr}$ \\
  \opcode{ldelem}    & Objects    & $\mathit{arr}, i \to v$ \\
  \opcode{stelem}    & Objects    & $\mathit{arr}, i, v \to$ \\
  \opcode{call}      & Calls      & $\mathit{args}\ldots \to [r]$ \\
  \opcode{callvirt}  & Calls      & $\mathit{args}\ldots, \mathit{obj} \to [r]$ \\
  \opcode{ret}       & Calls      & $[v] \to$ (returns to caller) \\
  \opcode{if}        & Control    & (condition-dependent) \\
  \opcode{while}     & Control    & (condition-dependent) \\
  \opcode{break}     & Control    & --- \\
  \opcode{continue}  & Control    & --- \\
  \opcode{try}       & Control    & --- \\
  \opcode{throw}     & Control    & $v \to$ (raises exception) \\
  \bottomrule
\end{longtable}

\end{document}

